More on the Slavnov products of spin chains, KP hierarchy tau and Baker-Akhiezer functions

### 2025-05-23 draft.tex
finish paper

\begin{abstract}
Connections between classical and quantum integrable systems are
analyzed from the viewpoint of the Slavnov products of Bethe
states. It is well known that, modulo model-dependent aspects, the
functional structure of the Slavnov products takes the form of
determinants. Using new results on the general structure of rational
and trigonometric models, we establish that, as long as the model
satisfies certain conditions, its Slavnov product is a tau function of
the KP hierarchy, thus generalizing some known results. We also show
that the Slavnov products can be expanded in terms of other tau
functions, and we prove that their homogeneous limit can be calculated
as a Wronskian of functions related to the eigenvalues of the transfer
matrices. Finally, we calculate the Baker–Akhiezer functions
associated with these Slavnov products, and show that, modulo a
multiplicative universal term, the Baker–Akhiezer functions have a
closed determinantal formula.
\end{abstract}

@@

\setcounter{tocdepth}{2}
\tableofcontents

@@

The main ingredients for the Slavnov product are the Bethe
wavefunctions themselves. As such, the specific details of these
objects depend on the model, symmetries and their representations, and
on the boundary conditions. Nevertheless, the general
\textbf{functional structure} of Slavnov products across radically
different models is basically the same. More specifically, modulo some
multiplicative factors, all these expressions take the form of
determinants, and it is this basic property that allows us to
manipulate these objects and prove that they are KP and/or Toda tau
functions.

@@

In Section 2, we review the main aspects of the
work~\cite{Belliard:2019bfz}. This section also fixes some of the
notation we use in the remainder of this work and in future
investigations. Section 3 presents two new results. First, we rewrite
relevant formulas established in~\cite{Belliard:2019bfz} and prove
that these Slavnov products are KP tau functions expressed in terms of
alternant determinants. We also discuss some contrasts between this
formula and certain results that have previously appeared in matrix
theory investigations.

These tau functions are written in terms of two sets of
parameters. The first set is given by arbitrary complex numbers, and
the second set is given by the Bethe roots of the spin chain.  In
Section 3, we also find a basis for the Slavnov products, and this
basis is itself formed by tau functions. The results of Section 3
allow us to investigate the case where the complex parameters are
close to the Bethe roots.

@@

Given that each function \(\bm{\tau}^{(\ell)}(\bm{z}^{(\ell)}, \bm{v})\), for
\(\ell = 1, \dots, n+1\), is a distinct but related tau function, one
can define a vector
\begin{equation}
  \bm{T}(\bm{Z}, \bm{v}) = 
\begin{bmatrix}
\bm{\tau}^{(1)}(\bm{z}^{(1)}, \bm{v}) \\ 
\bm{\tau}^{(2)}(\bm{z}^{(2)}, \bm{v}) \\ 
\vdots \\ 
\bm{\tau}^{(n+1)}(\bm{z}^{(n+1)}, \bm{v})
\end{bmatrix}
\end{equation}
where \(\bm{Z} = (\bm{z}^{(1)}, \bm{z}^{(2)}, \dots,
\bm{z}^{(n+1)})\). This is a vector of tau functions of the KP
hierarchy. This observation suggests that there is, in fact, an
underlying multicomponent KP hierarchy in this context, and that the
linear system~(\ref{eq:linear-system}), which we can write as
\begin{equation}
\left( \bm{L} - \bm{\Lambda} \right) \bm{T} = \bm{0},
\end{equation}
defines a reduction of this multicomponent KP hierarchy. This
reduction ultimately describes constraints that relate the different
sets of parameters \(\bm{z}^{(\ell)}\).  We have not explored this
idea in this work, but we hope to return to this problem in future
publications.

@@

In this paper we have discussed some properties of the Slavnov
products and their relation with the tau functions of the KP
hierarchy.

An important initial observation is that the general form of this tau
function is written in terms of some alternant matrix, and this form
is known since (cite papers on the matrix models and so on). But the
functions associated to the current paper is drastically different
from those papers, and do not expand a point in the
Grassmannian. Therefore, it is necessary to better understand some of
these properties. In this paper we start exploring these results.

....

We have also started our exploration of the Baker-Akhiezer function since it seems to
be the relevant object of our investigation.

Some future interesting directions. We might want to calculate the
solutions of the KP equation to see what they really mean. Perhaps
consider analitic and numerical calculations, since the calculation of
the determinants become complicated even for small values of \(n\).

More interestingly, one may try to consider the thermodynamic limit of
the Slavnov product and explore what it means from the viewpoint of
the Baker-Akhiezer function.  See the Limit \(L \to \infty\) and \(N
\to \infty\) but sufficiently slower than \(L\).  I think that the
matrix \(\mathcal{M}\) is a trace class operator and we write the BA
function as a Fredholm determinant.

I think that this BA function will be similar to the soliton-like
solution discussed in section 3.6.2 (page 41) of~\cite{Zabrodin2018}.
My functions seem to have a pole of order \(N\) at infinity. And \(N\)
zeros at the Bethe roots.

Finally, we can explore the elliptic case and see if one can see if
there exists any similar relation on the elliptic case.  Elliptic case
- see slavnov, zabrodin's paper

\subsubsection*{Acknowledgments}
I would like to thank...

>>>>

\begin{abstract}
 Connections between classical and quantum integrable systems are
 analyzed from the viewpoint of Slavnov products of Bethe states. It
 is well known that, modulo model dependent aspects, the functional
 structure of Slavnov products generally takes the form of
 determinants. Building on recent results on the structure of rational
 and trigonometric models, we show that, provided certain conditions
 are satisfied, the Slavnov product of a given model can be
 interpreted as a tau function of the KP hierarchy, thus extending
 known results in a more general setting. Moreover, we show that
 Slavnov products can be expanded in terms of other tau functions. We
 also prove that their homogeneous limit can be systematically
 expressed as a Wronskian of functions related to the eigenvalues of
 the transfer matrices.  Finally, we compute the Baker–Akhiezer
 functions associated with these Slavnov products and show that, apart
 from a universal multiplicative factor, they admit a closed
 determinantal representation.
\end{abstract}

@@

\setcounter{tocdepth}{1}
\tableofcontents

@@

The main ingredients for the Slavnov product are the Bethe
wavefunctions themselves. As such, the specific details of these
objects depend on the model, symmetries and their representations, and
on the boundary conditions. Nevertheless, the general
functional structure of Slavnov products across radically
different models is basically the same. More specifically, modulo some
multiplicative factors, all these expressions take the form of
determinants, and it is this basic property that allows us to
manipulate these objects and prove that they are KP and/or Toda tau
functions.

@@

Let us now provide a brief overview of the content and structure of
our work. In Section 2, we offer a review of the main findings
presented in the work of Belliard and
Slavnov~\cite{Belliard:2019bfz}. In addition, we use this section to
carefully establish and clarify the notation that will be employed
throughout the remainder of this paper, as well as in related future
investigations.

Section 3 presents two new results and a conjecture. First, we rewrite
relevant formulas established in~\cite{Belliard:2019bfz} and prove
that these Slavnov products are KP tau functions expressed in terms of
alternant determinants. We also discuss some contrasts between this
formula and certain results that have previously appeared in matrix
theory investigations.  These tau functions are written in terms of
two sets of parameters. The first set consists of arbitrary complex
numbers, and the second set consists of the Bethe roots of the spin
chain. In Section 3, we also find a basis for the Slavnov products,
and this basis is itself formed by tau functions. The results of
Section 3 allow us to investigate the case where the complex
parameters are close to the Bethe roots.

In this section, we also conjecture (and only present some general
evidence) that these tau functions indicate that we are, in fact,
dealing with a multicomponent KP hierarchy, and that the linear
equations discussed by Belliard and Slavnov~\cite{Belliard:2019bfz}
describe a reduction of this larger integrable hierarchy.

@@

Given that each function \(\bm{\tau}^{(\ell)}(\bm{z}^{(\ell)}, \bm{v})\), for
\(\ell = 1, \dots, n+1\), is a distinct but related tau function, one
can define a vector
\begin{equation}
  \bm{T}(\bm{Z}, \bm{v}) = 
\begin{bmatrix}
\bm{\tau}^{(1)}(\bm{z}^{(1)}, \bm{v}) \\ 
\bm{\tau}^{(2)}(\bm{z}^{(2)}, \bm{v}) \\ 
\vdots \\ 
\bm{\tau}^{(n+1)}(\bm{z}^{(n+1)}, \bm{v})
\end{bmatrix}
\end{equation}
where \(\bm{Z} = (\bm{z}^{(1)}, \bm{z}^{(2)}, \dots,
\bm{z}^{(n+1)})\). This is a vector of tau functions of the KP
hierarchy.

\begin{conjecture}
This observation suggests that there is, in fact, an
underlying multicomponent KP hierarchy in this context, and that the
linear system~(\ref{eq:linear-system}), which we can write as
\begin{equation}
\left( \bm{L} - \bm{\Lambda} \right) \bm{T} = \bm{0},
\end{equation}
defines a reduction of this multicomponent KP hierarchy. This
reduction ultimately describes constraints that relate the different
sets of parameters \(\bm{z}^{(\ell)}\).  We have not explored this
idea in this work, but we hope to return to this problem in future
publications.
\end{conjecture}

@@

In this work, we have discussed several properties of the Slavnov
products arising in quantum integrable models and their deep
connection with the tau functions of the KP hierarchy. We can
summarize our main findings as follows. Our initial result
demonstrates that the general structure of these tau functions can be
expressed in terms of an alternant matrix. This general formulation
firmly establishes the identification of the Slavnov product as a tau
function within the framework of the KP classical integrable hierarchy.

We have also proved that these Slavnov products admit a basis
expansion in terms of other tau functions, thereby establishing that
we are dealing with a particularly distinguished object within this
framework. Moreover, we have discussed a conjecture suggesting the
existence of a multicomponent KP hierarchy underlying all of our
results.

The behaviour of these functions near the Bethe roots of the quantum
integrable systems, as well as the homogeneous limit of the Slavnov
product, has also been discussed. Finally, we conclude our work with a
brief analysis of the Baker-Akhiezer function. The most important
result for us is that we have shown that the Baker-Akhiezer function,
modulo a universal multiplicative factor, also admits a determinantal
form.

Evidently, there are many results that can be extended in this work,
and several aspects that deserve further investigation. Let us list
some of these problems, ranging from relatively straightforward
applications to more substantial challenges.

A simple problem to be discussed is the physical meaning of the
solutions of the KP equation that can be constructed using the tau
functions explored in this work. Perhaps one might consider both
analytical and numerical approaches, since the calculation of the
determinants becomes complicated even for relatively small values of
\(n\).

Another important aspect of this work is to achieve a better
understanding of the Baker-Akhiezer functions associated with the
Slavnov products. Here, we have only scratched the surface of these
objects, and a full description of their properties is still
lacking. We are currently investigating the thermodynamic limit of
this system. In particular, we are studying the Slavnov product and
its interpretation from the viewpoint of the Baker-Akhiezer
function. For example, how to properly describe the limit \(L \to
\infty\) and \(N \to \infty\), with \(N\) growing sufficiently
slower than \(L\). I believe that the matrix \(\mathcal{M}\) is a
trace-class operator, allowing us to express the Baker-Akhiezer
function as a Fredholm determinant. We hope to report new results
on this line of investigation soon.

Finally, there is a more challenging problem to be addressed: the
description of the elliptic case. While some determinantal formulas
are known for the partially on-shell scalar product of Bethe states,
it is not yet clear whether these objects are also related to tau
functions of integrable hierarchies.  This characterization is another
problem we are currently investigating, and we hope to have some
results to report in the future.

%\subsubsection*{Acknowledgments}
%I would like to thank the Fluminense Federal University for
